class: inverse, center, middle

# Tidy data
---

## Validate
```{r}
data(cars)
head(cars)

library(validate)
rules <- validator(speed >= 0, 
                   dist >= 0,
                   speed/dist <= 1.5)

out <- confront(cars, rules)
summary(out)
```

---
```{r}
violating(cars, out)
```

---
## Clean data
*This section has been taken from: Grolemund, G., & Wickham, H. (2017). R for Data Science. O’Reilly Media.*
```{r}
library(tidyr)

head(who[,1:10])
```

WHO: A subset of data from the World Health Organization Global Tuberculosis Report. It is a data frame with 7,240 rows and the columns: 1) `country`: Country name 2)`iso2`, `iso3`: 2 & 3 letter ISO country codes 3) `year`:Year

---
```{r}
head(who[,5:12])
```

 4) `new_sp_m014` - `new_rel_f65`: Counts of new TB cases recorded by group. Column names encode three variables that describe the group:    
 
  * The first three letters of each column denote whether the column contains `new` or `old` cases of TB. In this dataset, each column contains new cases.
  * The next two letters describe the type of TB:
     * `rel` stands for cases of relapse
     * `ep` stands for cases of extrapulmonary TB
     * `sn` stands for cases of pulmonary TB that could not be diagnosed by a pulmonary smear (smear negative)
     * `sp` stands for cases of pulmonary TB that could be diagnosed by a pulmonary smear (smear positive)
     * The sixth letter gives the sex of TB patients. The dataset groups cases by males (`m`) and females (`f`).
---
```{r}
who1 <- 
  who %>% 
  pivot_longer(
    cols = -c(country, iso2, iso3, year), 
    names_to = "key", 
    values_to = "cases", 
    values_drop_na = TRUE
  )

who1

```

---
.pull-left[
```{r}
library(dplyr)
who1 %>%
  count(key)
```
]
.pull-left[
```{r}
who1 %>%
  count(key) %>%
  arrange(n)
```
]

---
.pull-left[
```{r}
library(stringr)
who2 <- 
  who1 %>% 
  mutate(key = str_replace(key,
                           "newrel", 
                           "new_rel"))

who2 %>%
  count(key) %>%
  arrange(n)
```
]

.pull-right[
```{r}
who3 <- 
  who2 %>% 
  separate(key, 
           c("new", "type", "sexage"), 
           sep = "_") %>%
  select(-iso2, -iso3)

who3
```
]

---
```{r}
who3 %>%
  count(new)
```

```{r}
who4 <- 
  who3 %>% 
  separate(sexage, c("sex", "age"), sep = 1)

who4
```

---
```{r}
who %>%
  pivot_longer(
    cols = new_sp_m014:newrel_f65, 
    names_to = "key", 
    values_to = "cases", 
    values_drop_na = TRUE
  ) %>% 
  mutate(
    key = stringr::str_replace(key, "newrel", "new_rel")
  ) %>%
  separate(key, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1)
```