
class: inverse, center, middle

# Communicate
---

## R Markdown

R Markdown provides a framework combining code, its results, and text. R Markdown documents are fully reproducible and support dozens of output formats (e.g. PDFs, Word files, Powerpoint, HTML, etc).

This is an R Markdown file, a plain text file that has the extension `.Rmd`

.pull-left[
```{r echo = FALSE, comment = ""}
cat(htmltools::includeText(here::here("rmd", "presentation", "rmarkdown", "sample1.Rmd")))
```
]

.pull-right[
```{r echo = FALSE, comment = ""}
cat(htmltools::includeText(here::here("rmd", "presentation", "rmarkdown", "sample2.Rmd")))
```
]

---
It contains three important types of content:

1.  A **header** surrounded by `---`  
2.  **Chunks** of R code surrounded by ```` ``` ````  
3.  Text mixed with inline code ``` `r ` ```  

.pull-left[
```{r echo = FALSE, comment = ""}
cat(htmltools::includeText(here::here("rmd", "presentation", "rmarkdown", "sample1.Rmd")))
```
]

.pull-right[
```{r echo = FALSE, comment = ""}
cat(htmltools::includeText(here::here("rmd", "presentation", "rmarkdown", "sample2.Rmd")))
```
]

---
### The output is:

```{r echo = FALSE, child = here::here("rmd", "presentation", "rmarkdown", "sample.Rmd")}
```

---
### Parameters

* R Markdown documents can include one or more parameters. 

* Parameters are useful when you want to regenerate the same report with different inputs

* This is very hadny when you want to generate site reports (i.e. same reports for different sites)

.pull-left[
```{r echo = FALSE, comment = ""}
cat(htmltools::includeText(here::here("rmd", "presentation", "rmarkdown", "sample1_params.Rmd")))
```
]

.pull-right[
```{r echo = FALSE, comment = ""}
cat(htmltools::includeText(here::here("rmd", "presentation", "rmarkdown", "sample2_params.Rmd")))
```
]

---
### Making several reports
```{r, echo = TRUE, eval = FALSE}
library(knitr)

render(input = "path_to_the_Rmd_file",
       params = "list_of_the_parameters_to_use",
       output_format = "output_fromat (e.g. pdf_document)",
       output_file = "the_name_of_the_output_file",
       ...)

make_my_report <- function(SiteID) {
  render(input = "report.Rmd",
         params = list(SiteID = SiteID),
         output_format = "pdf_document"
         output_file = glue::glue("report_{SiteID}_{Sys.Date()}.pdf"))
}

input_df <- data.frame(SiteID = c("group_1", "group_2", "etc"))

purrr::pwalk(input_df, make_my_report)
```

---
### R Markdown formats

* HTML
* PDF
* WORD
* RTF
* POWEPOINT
* HTML Dashboards (requires Flexdashboard)
* Interactive Dashboards (requires Shiny)
* Websites (needs a little additional infrastructure)
* ETC

---
## Interactive outputs and animation

```{r}
DT::datatable(head(registryr::fake_data), fillContainer = FALSE, options = list(pageLength = 4))
```

---
```{r}
graph <-
  registryr::fake_data %>%
  group_by(site_id) %>%
  summarise(mortality_rate = sum(dead) / n()) %>%
  mutate(n = seq(100, 2000, length.out = 20)) %>%
  ggplot(aes(x = n, y = mortality_rate)) +
  geom_point() +
  theme_bw()

plotly::ggplotly(graph)
```

---
```{r}
p <-
  registryr::fake_data %>%
  group_by(site_id) %>%
  summarise(mortality_rate = sum(dead) / n()) %>%
  mutate(year = seq(2001, 2020)) %>%
  ggplot(aes(x = year, y = mortality_rate)) +
  geom_line() +
  theme_bw()

p
```

---
```{r}
p + gganimate::transition_reveal(year)
```

---
## R Shiny
Shiny is an R package that makes it possible to build interactive web apps straight from R.

```{r, echo = FALSE}
knitr::include_app("https://vac-lshtm.shinyapps.io/ncov_tracker",
                   height = "500px")

```

---
```{r, echo = FALSE}
knitr::include_app("https://monash-corrp.shinyapps.io/DCQR_NNT_dashboard",
                   height = "600px")
```
